---
title: "PSA purification"
date: "`r Sys.Date()`"
---

```{r knitr-init, echo=F}
library(knitr)
## Global options
options(digits = 4)
opts_chunk$set(echo=F, cache=T, prompt=F, tidy=T, comment=NA, message=F, warning=F, dev="png", dev.args=list(type="cairo"), dpi=300)
```

```{r packages, results='hide'}
.packages <- c("DT")
.installed <- .packages %in% installed.packages()
if (length(.packages[!.installed])>0) install.packages(.packages[!.installed])
lapply(.packages, require, character.only=T)
```

<!-- # Authors -->

<!-- - Mercedes de Frutos Gómez ([mfrutos@iqog.csic.es](mailto:mfrutos@iqog.csic.es)) -->
<!-- - Alfredo Sánchez Alberca ([asalber@ceu.es](mailto:asalber@ceu.es)) -->

# Summary

Prostate-Specific Antigen (PSA) is usually purified before applying capillary electrophoresis.

The purpose of this study is to determine if purification significantly alters the sprectrum of PSA and if there are significant differences among six different purification methods:
<ol type="A">
<li>antiPSA column</li>
<li>IgG1 first column</li>
<li>IgG1 second column</li>
<li>IgG1 two columns</li>
<li>Off-line coupled columns</li>
<li>On-line coupled columns</li>
</ol>

# Objetives

- To determine if purification significantly alters the sprectrum of PSA.
- Compare six different purification methods.

# Type of study 

Analityc observational cross-sectional and comparative study.

<!-- # Inclusion criteria -->

<!-- Individuals with prostate cancer or benign prostatic hyperplasia. -->

<!-- # Variables -->

<!-- ```{r variables} -->
<!-- vars <- read.csv(file = "data/variables.csv", header = T, sep = ",") -->
<!-- datatable(vars, rownames = F, escape=F, class = 'display', options = list(dom="ltip")) -->
<!-- ``` -->

# Sample size

```{r data-loading}
source("data-loading.R")
```


Total number of analysis: `r nrow(data)/10`
```{r sample-size-groups}
countTable <- tbl_df(data) %>% count(Method)
countTable$n <- countTable$n / 10
pander(countTable)
```

# Glosary

- PSA: Prostate-Specific Antigen

<!--chapter:end:index.Rmd-->

---
title: "Descriptive statistics"
date: "`r Sys.Date()`"
---

```{r knitr-init, echo=F}
library(knitr)
## Global options
options(digits = 4)
opts_chunk$set(echo=F, cache=T, prompt=F, tidy=T, comment=NA, message=F, warning=F, dev="png", dev.args=list(type="cairo"), dpi=300)
color1 <- "#56B4E9"
color2 <- "#CC79A7"
```

```{r data-loading, results='hide'}
source("data-loading.R")
```


# Non-pufied aliquot

```{r descriptive-statistics-non-purified}
statistics <- data.non.purified %>% 
  group_by(Peak) %>% 
  summarise(n = n(), mean = mean(Area, na.rm = T), sd = sd(Area, na.rm = T)) %>%
  mutate(se = sd/sqrt(n), lower.ci = mean-qt(1-(0.05/2), n-1)*se, upper.ci = mean+qt(1-(0.05/2), n-1)*se)
pander(statistics, split.cells = 1)
``` 

```{r scatter-plot-non-purified}
ggplot(data.non.purified, aes(x = Peak, y = Area)) + 
  geom_jitter(width = 0.2, colour = color1) + 
  geom_pointrange(data = statistics, mapping=aes(x=Peak, y=mean, ymin=mean-sd, ymax=mean+sd), shape = 95, size = 1.5, colour = color2 ) +
  ggtitle("Percentage of area covered by peaks of PSA for non-purified aliquot") +
  ylab("% Area") +
  theme(plot.title = element_text(hjust = 0.5))
```

```{r descriptive-statistics-groups-non-purified}
statistics <- data.non.purified %>% 
  group_by(Peak, Method) %>% 
  summarise(n = n(), mean = mean(Area, na.rm = T), sd = sd(Area, na.rm = T)) %>%
  mutate(se = sd/sqrt(n), lower.ci = mean-qt(1-(0.05/2), n-1)*se, upper.ci = mean+qt(1-(0.05/2), n-1)*se)
pander(statistics, split.cells = 1)
``` 

```{r scatter-plot-groups-non-purified}
ggplot(data.non.purified, aes(x = Peak, y = Area, colour = Method)) + 
  geom_point(position = position_dodge(width = 0.8)) + 
  geom_pointrange(data = statistics, mapping=aes(x=Peak, y=mean, colour = Method, ymin=mean-sd, ymax=mean+sd), position = position_dodge(width = 0.8), shape = 95, size = 1.5) +
  ggtitle("Percentage of area covered by peaks of PSA for non-purified aliquot") +
  ylab("% Area") +
  theme(plot.title = element_text(hjust = 0.5))
```

# Means and standard deviations

```{r scatter-plot}
ggplot(data, aes(x = Peak, y = Variation, colour=Method)) + 
  geom_point(position=position_dodge(width=0.5)) + 
  #stat_summary(fun.y="mean", size=3, shape=1, geom="point", position=position_dodge(width=0.5)) +
  ggtitle("Variation of percentage of area covered by peaks of PSA (before - after purification)") +
  ylab("% Area before - % Area after") +
  theme(plot.title = element_text(hjust = 0.5))
```

```{r descriptive-statistics-groups}
statistics <- data %>% group_by(Peak, Method) %>% 
  summarise(n = n(), mean = mean(Variation, na.rm = T), sd = sd(After, na.rm = T)) %>%
    mutate(se = sd/sqrt(n), lower.ci = mean-qt(1-(0.05/2), n-1)*se, upper.ci = mean+qt(1-(0.05/2), n-1)*se)
pander(statistics, split.cells = 1)
``` 

```{r confidence-intervals-means}
ggplot(data = statistics, aes(x = Peak, y = mean, colour = Method, shape = Method)) +
  geom_pointrange(aes(ymin = lower.ci, ymax = upper.ci), position = position_dodge(width = 0.5)) +
  ggtitle("Confidence intervals for the means of \n variation of percentage of areas covered by peaks of PSA") +
   ylab("% Area before - % Area after") +
  theme(plot.title = element_text(hjust = 0.5))
```

We can observe that the method with more variability are A, E and F and the ones with less variability C and D.

Intervals not containing 0 show a significant differences between the reference and the purified sample.

In general, method with less variability are preferable, even if there is a significant variation, when the variation is close to zero.

<!--chapter:end:descriptive.Rmd-->

---
title: "Method comparison"
date: "`r Sys.Date()`"
---

```{r knitr-init, echo=F}
library(knitr)
## Global options
options(digits = 4)
opts_chunk$set(echo=F, cache=T, prompt=F, tidy=T, comment=NA, message=F, warning=F, dev="png", dev.args=list(type="cairo"), dpi=300)
```

```{r data-loading, results='hide'}
source("data-loading.R")
```

# Comparison of means of methods E and F

```{r t-test-EF}
data.EF <- data[data$Method %in% c("E", "F"), ]
result <- data.EF %>% group_by(Peak, Method) %>%
  summarize(Variation = list(Variation)) %>%
  spread(Method, Variation) %>%
  mutate(t_value = t.test(unlist(E), unlist(F), var.equal = T)$statistic, 
         p_value = t.test(unlist(E), unlist(F), var.equal = T)$p.value,
         lower.ci = t.test(unlist(E), unlist(F), var.equal = T)$conf.int[1],
         upper.ci = t.test(unlist(E), unlist(F), var.equal = T)$conf.int[2])
pander(result[, c(1, 4:7)])
```

We can observe that the only peaks wehre there is a significant difference are peak 2 and 5. 

```{r confidence-intervals-EF}
statistics <- data.EF %>% group_by(Peak, Method) %>% 
  summarise(n = n(), mean = mean(Variation, na.rm = T), sd = sd(After, na.rm = T)) %>%
    mutate(se = sd/sqrt(n), lower.ci = mean-qt(1-(0.05/2), n-1)*se, upper.ci = mean+qt(1-(0.05/2), n-1)*se)
ggplot(data = statistics, aes(x = Peak, y = mean, colour = Method, shape = Method)) +
  geom_pointrange(aes(ymin = lower.ci, ymax = upper.ci), position = position_dodge(width = 0.5)) +
  ggtitle("Confidence intervals for the means of \n variation of percentage of areas covered by peaks of PSA") +
   ylab("% Area before - % Area after") +
  theme(plot.title = element_text(hjust = 0.5))
```


<!--chapter:end:method-comparison.Rmd-->

