---
title: "Descriptive statistics"
date: "`r Sys.Date()`"
---

```{r knitr-init, echo=F}
library(knitr)
## Global options
options(digits = 4)
opts_chunk$set(echo=F, cache=T, prompt=F, tidy=T, comment=NA, message=F, warning=F, dev="png", dev.args=list(type="cairo"), dpi=300)
```

```{r data-loading, results='hide'}
source("data-loading.R")
```

# Means and standard deviations

```{r scatter-plot}
ggplot(data, aes(x = Peak, y = Variation, colour=Method)) + 
  geom_point(position=position_dodge(width=0.5)) + 
  #stat_summary(fun.y="mean", size=3, shape=1, geom="point", position=position_dodge(width=0.5)) +
  ggtitle("Variation of percentage of area covered by peaks of PSA (before - after purification)") +
  ylab("% Area before - % Area after") +
  theme(plot.title = element_text(hjust = 0.5))
```

```{r descriptive-statistics-groups}
statistics <- data %>% group_by(Peak, Method) %>% 
  summarise(n = n(), mean = mean(Variation, na.rm = T), sd = sd(After, na.rm = T)) %>%
    mutate(se = sd/sqrt(n), lower.ci = mean-qt(1-(0.05/2), n-1)*se, upper.ci = mean+qt(1-(0.05/2), n-1)*se)
pander(statistics, split.cells = 1)
``` 

```{r confidence-intervals-means}
ggplot(data = statistics, aes(x = Peak, y = mean, colour = Method, shape = Method)) +
  geom_pointrange(aes(ymin = lower.ci, ymax = upper.ci), position = position_dodge(width = 0.5)) +
  ggtitle("Confidence intervals for the means of \n variation of percentage of areas covered by peaks of PSA") +
   ylab("% Area before - % Area after") +
  theme(plot.title = element_text(hjust = 0.5))
```

We can observe that the method with more variability are A, E and F and the ones with less variability C and D.

Intervals not containing 0 show a significant differences between the reference and the purified sample.

In general, method with less variability are preferable, even if there is a significant variation, when the variation is close to zero.
