---
title: "PSA purification"
date: "2019/09/09"
---

```{r knitr-init, echo=F}
library(knitr)
## Global options
options(digits = 4)
opts_chunk$set(echo=F, cache=T, prompt=F, tidy=T, comment=NA, message=F, warning=F, dev="png", dev.args=list(type="cairo"), dpi=300)
```

```{r packages, results='hide'}
.packages <- c("tidyverse", "kableExtra")
.installed <- .packages %in% installed.packages()
if (length(.packages[!.installed])>0) install.packages(.packages[!.installed])
lapply(.packages, require, character.only=T)
```

<!-- # Authors -->

<!-- - Mercedes de Frutos Gómez ([mfrutos@iqog.csic.es](mailto:mfrutos@iqog.csic.es)) -->
<!-- - Alfredo Sánchez Alberca ([asalber@ceu.es](mailto:asalber@ceu.es)) -->

# Summary

Prostate-Specific Antigen (PSA) is usually purified before applying capillary electrophoresis.

The purpose of this study is to determine if purification significantly alters the sprectrum of PSA and if there are significant differences among six different purification methods:
<ol type="A">
<li>antiPSA column</li>
<li>IgG1 first column</li>
<li>IgG1 second column</li>
<li>IgG1 two columns</li>
<li>Off-line coupled columns</li>
<li>On-line coupled columns</li>
</ol>

# Objetives

- To determine if purification significantly alters the sprectrum of PSA.
- Compare six different purification methods.

# Type of study 

Analityc observational cross-sectional and comparative study.

<!-- # Inclusion criteria -->

<!-- Individuals with prostate cancer or benign prostatic hyperplasia. -->

<!-- # Variables -->

<!-- ```{r variables} -->
<!-- vars <- read.csv(file = "data/variables.csv", header = T, sep = ",") -->
<!-- datatable(vars, rownames = F, escape=F, class = 'display', options = list(dom="ltip")) -->
<!-- ``` -->

# Sample size

```{r data-loading}
source("data-loading.R")
```


Total number of analysis: `r nrow(data)/10`
```{r sample-size-groups}
data %>%
  group_by(Method, Type) %>%
  count() %>%
  mutate(n = n/10) %>%
  kable() %>%
  kable_styling(bootstrap_options = "hover", full_width = F)
```

# Glosary

- PSA: Prostate-Specific Antigen