---
title: "Method comparison"
date: "`r Sys.Date()`"
---

```{r knitr-init, echo=F}
library(knitr)
## Global options
options(digits = 4)
opts_chunk$set(echo=F, cache=T, prompt=F, tidy=T, comment=NA, message=F, warning=F, dev="png", dev.args=list(type="cairo"), dpi=300)
color1 <- "#F8766D"
```

```{r packages, results='hide'}
# Load packages
.packages <- c("tidyverse", "rstatix", "crosstalk", "reactable", "htmltools", "ggpubr", "kableExtra")
.installed <- .packages %in% installed.packages()
if (length(.packages[!.installed])>0) install.packages(.packages[!.installed])
lapply(.packages, require, character.only=T)
```

```{r data-loading, results='hide'}
source("data-loading.R")
```

## Confidence intervals for means

```{r ci-means}
stats <- data %>%
  group_by(Peak, Method, Type) %>%
  get_summary_stats(Area, type = "mean_ci") %>%
  mutate(conf.int. = paste0("(", mean - ci, ", ", mean + ci, ")"))
 
stats2 <- SharedData$new(select(stats, -c("variable", "ci")))

bscols(
  widths = c(2, 10),
  list(
    tags$h3("Filter"),
    filter_select("Method", "Method", stats2, ~Method),
    filter_select("Type", "Type", stats2, ~Type),
    filter_select("Peak", "Peak", stats2, ~Peak)
  ),
  reactable(stats2, minRows = 10)
)
```

```{r ci-groups-faceted}
ggplot(stats) + 
  facet_wrap(vars(Peak), scales = "free") +
  geom_pointrange(mapping = aes(x = Method, y = mean, ymin = mean-ci, ymax = mean+ci, colour = Type), size = 0.3, position = position_dodge(width = 0.7)) +
  ggtitle("95% confidence intervals for the mean of covered area by peaks") +
  xlab("") +
  ylab("% Area") +
  theme(plot.title = element_text(hjust = 0.5)) +
  theme_bw()
```

## Confidence intervals for the difference of means of non-purified and purified aliquots

```{r ci-diff-means}
stats.diff <- data %>%
  group_by(Peak, Method) %>%
  t_test(Area ~ Type, paired = F, detailed = T) %>%
  mutate(Diff.means = round(estimate1 - estimate2, 3), Mean.non.purified = round(estimate1, 3), Mean.purified = round(estimate2, 3), conf.int = paste0("(", round(conf.low, 3), ", ", round(conf.high, 3), ")"))

stats.diff2 <- SharedData$new(select(stats.diff, Method, Peak, Mean.non.purified, Mean.purified, conf.int))

bscols(
  widths = c(2, 10),
  list(
    tags$h3("Filter"),
    filter_select("Method", "Method", stats.diff2, ~Method),
    filter_select("Peak", "Peak", stats.diff2, ~Peak)
  ),
  reactable(stats.diff2, minRows = 10)
)
```

```{r ci-diff-groups-faceted}
ggplot(stats.diff) + 
  facet_wrap(vars(Peak), scales = "free") +
  geom_pointrange(mapping = aes(x = Method, y = Diff.means, ymin = conf.low, ymax = conf.high), colour = color1, size = 0.3, position = position_dodge(width = 0.7)) +
  ggtitle("95% confidence intervals for the difference of non-purified and purified means of covered area by peaks") +
  xlab("") +
  ylab("% Area") +
  theme(plot.title = element_text(hjust = 0.5)) +
  theme_bw()
```


## Comparison of non-purified and purified aliquots

As the sample size per group is very small (from 3 to 6) instead of using a t-test we use the Mann-Whitney test, a non-parametric method to compare the ranks.

### Method A
```{r mann-withney-A}
data %>%
  filter(Method == "A") %>%
  group_by(Peak) %>%
  wilcox_test(Area ~ Type, paired = F) %>%
  select(-.y.) %>%
  kable()  %>%
  kable_styling(bootstrap_options = "hover", full_width = F)
```

Thus, there are significant differences between the areas of peaks 1, 3 and 10.

```{r boxplot-A}
data %>% 
  filter(Method == "A") %>%
  ggplot(aes(x = Type, y = Area, col = Type)) +
  facet_wrap(vars(Peak), scales = "free") +
  geom_boxplot() +
  guides(col = "none") +
  ggtitle("Boxplot of area covered by peaks in non-purified and purified aliquots with method A")
```


### Method B
```{r mann-withney-B}
data %>%
  filter(Method == "B") %>%
  group_by(Peak) %>%
  wilcox_test(Area ~ Type, paired = F) %>%
  select(-.y.) %>%
  kable()  %>%
  kable_styling(bootstrap_options = "hover", full_width = F)
```

Thus, there are no significant differences between the areas for all the peaks.

```{r boxplot-B}
data %>% 
  filter(Method == "B") %>%
  ggplot(aes(x = Type, y = Area, col = Type)) +
  facet_wrap(vars(Peak), scales = "free") +
  geom_boxplot() +
  guides(col = "none") +
  ggtitle("Boxplot of area covered by peaks in non-purified and purified aliquots with method B")
```

### Method C
```{r mann-withney-C}
data %>%
  filter(Method == "C") %>%
  group_by(Peak) %>%
  wilcox_test(Area ~ Type, paired = F) %>%
  select(-.y.) %>%
  kable()  %>%
  kable_styling(bootstrap_options = "hover", full_width = F)
```

Thus, there are no significant differences between the areas for all the peaks.

```{r boxplot-C}
data %>% 
  filter(Method == "C") %>%
  ggplot(aes(x = Type, y = Area, col = Type)) +
  facet_wrap(vars(Peak), scales = "free") +
  geom_boxplot() +
  guides(col = "none") +
  ggtitle("Boxplot of area covered by peaks in non-purified and purified aliquots with method C")
```

### Method D
```{r mann-withney-D}
data %>%
  filter(Method == "D") %>%
  group_by(Peak) %>%
  wilcox_test(Area ~ Type, paired = F) %>%
  select(-.y.) %>%
  kable()  %>%
  kable_styling(bootstrap_options = "hover", full_width = F)
```

Thus, there are no significant differences between the areas for all the peaks.

```{r boxplot-D}
data %>% 
  filter(Method == "D") %>%
  ggplot(aes(x = Type, y = Area, col = Type)) +
  facet_wrap(vars(Peak), scales = "free") +
  geom_boxplot() +
  guides(col = "none") +
  ggtitle("Boxplot of area covered by peaks in non-purified and purified aliquots with method D")
```

### Method E
```{r mann-withney-E}
data %>%
  filter(Method == "E") %>%
  group_by(Peak) %>%
  wilcox_test(Area ~ Type, paired = F) %>%
  select(-.y.) %>%
  kable()  %>%
  kable_styling(bootstrap_options = "hover", full_width = F)
```

Thus, there are significant differences between areas of peaks 1, 2, 5, 7, 9 and 10.

```{r boxplot-E}
data %>% 
  filter(Method == "E") %>%
  ggplot(aes(x = Type, y = Area, col = Type)) +
  facet_wrap(vars(Peak), scales = "free") +
  geom_boxplot() +
  guides(col = "none") +
  ggtitle("Boxplot of area covered by peaks in non-purified and purified aliquots with method E")
```

### Method F
```{r mann-withney-F}
data %>%
  filter(Method == "F") %>%
  group_by(Peak) %>%
  wilcox_test(Area ~ Type, paired = F) %>%
  select(-.y.) %>%
  kable()  %>%
  kable_styling(bootstrap_options = "hover", full_width = F)
```

Thus, there are significant differences between areas of peaks 3, 4, 5 and 7.

```{r boxplot-F}
data %>% 
  filter(Method == "F") %>%
  ggplot(aes(x = Type, y = Area, col = Type)) +
  facet_wrap(vars(Peak), scales = "free") +
  geom_boxplot() +
  guides(col = "none") +
  ggtitle("Boxplot of area covered by peaks in non-purified and purified aliquots with method F")
```

## Comparison of non-purified aliquots

Again, as the sample size is very small (from 3 to 6) instead of using an ANOVA test we use the Kruskal-Wallis test, a non-parametric method to compare the ranks.

```{r kruskal-wallis-non-purified}
data.non.purified %>%
  group_by(Peak) %>%
  kruskal_test(Area ~ Method) %>%
  select(Peak, n, statistic, df, p) %>%
  kable()  %>%
  kable_styling(bootstrap_options = "hover", full_width = F)
```

We observe significant differences between at least two reference samples in peaks 1, 3, 5 and 8.

```{r dunn-non-purified}
stats <- data.non.purified %>%
  group_by(Peak) %>%
  dunn_test(Area ~ Method, p.adjust.method = "bonferroni") 

stats <- SharedData$new(select(stats, -.y., -statistic))

bscols(
  widths = c(2, 10),
  list(
    tags$h3("Filter"),
    filter_select("Peak", "Peak", stats, ~Peak)
  ),
  reactable(stats, minRows = 10,
            columns = list(
              p = colDef(format = colFormat(digits = 3)),
              p.adj = colDef(format = colFormat(digits = 3))
            )
  )
)
```

Comparing the referece samples by pairs we can observe:

- For peak 3, there are significant differences between the reference samples A and F and B and F.
- For peak 5, there are significant differences between the reference samples C and D.


## Comparison between purified aliquots with methods E and F

Again, as the sample size per group is very small (from 3 to 6) instead of using a t-test we use the Mann-Whitney test.

```{r mann-withney-E-F}
data.purified %>%
  filter(Method %in% c("E", "F")) %>%
  group_by(Peak) %>%
  wilcox_test(Area ~ Method, paired = F) %>%
  select(-.y.) %>%
  kable()  %>%
  kable_styling(bootstrap_options = "hover", full_width = F)
```

Thus, there is no significant difference between methods E and F.

```{r boxplot-E-F}
data.purified %>% 
  filter(Method %in% c("E", "F")) %>%
  ggplot(aes(x = Method, y = Area, col = Method)) +
  facet_wrap(vars(Peak), scales = "free") +
  geom_boxplot() +
  guides(col = "none") +
  ggtitle("Boxplot of area covered by peaks in purified aliquots with methods E and F")
```


## Comparison of purified aliquots (all methods)

Again, as the sample size is very small (from 3 to 6) instead of using an ANOVA test we use the Kruskal-Wallis test.

```{r kruskal-wallis-purified}
data.purified %>%
  group_by(Peak) %>%
  kruskal_test(Area ~ Method) %>%
  select(Peak, n, statistic, df, p) %>%
  kable()  %>%
  kable_styling(bootstrap_options = "hover", full_width = F)
```

We observe significant differences between at least two methods in peaks 1, 2, 5, 8, 9 and 10.

```{r dunn-purified}
stats <- data.purified %>%
  group_by(Peak) %>%
  dunn_test(Area ~ Method, p.adjust.method = "bonferroni") 

stats <- SharedData$new(select(stats, -.y., -statistic))

bscols(
  widths = c(2, 10),
  list(
    tags$h3("Filter"),
    filter_select("Peak", "Peak", stats, ~Peak)
  ),
  reactable(stats, minRows = 10,
            columns = list(
              p = colDef(format = colFormat(digits = 3)),
              p.adj = colDef(format = colFormat(digits = 3))
            )
  )
)
```

Comparing the methods by pairs we can observe that there are significant diferences between methods C and E only for peak 5.


## Protein recovery comparison (Methods E and F)

Once again we use the Mann-Whitney test due to the versy small sample sizes.

```{r mann-withney-recovery-E-F}
data.recovery %>%
  wilcox_test(Recovery ~ Method, paired = F) %>%
  select(-.y.) %>%
  kable()  %>%
  kable_styling(bootstrap_options = "hover", full_width = F)
```

Thus there is a significant difference for alpha = 0.1 but not for alpha = 0.05. 

```{r boxplot-recovery-E-F}
data.recovery %>% 
  ggplot(aes(x = Method, y = Recovery, col = Method)) +
  geom_boxplot() +
  guides(col = "none") +
  ggtitle("Boxplot of protein recovery with methods E and F")
```

Regardin the plot, it seems that there is a clear difference betwen the two methods but this difference is not significant for alpha = 0.05. Using t-test, that has more power we observe below a significant difference, however with such a small sample we can not guarantee the normality condition requiered by this test.

```{r t-test-E-F}
data.recovery %>%
  t_test(Recovery ~ Method, paired = F) %>%
  select(-.y.) %>%
  kable()  %>%
  kable_styling(bootstrap_options = "hover", full_width = F)
```


